Introduction to Structural Causal Models

### 2021-05-27 scm_short_presentation.Rmd
intermediate

@@

```{tikz,fig.cap="Structural Causal Model", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (E) {E};
\node[my node,right = 1 cm of E, fill=gray!30](C){C};
\node[latent,left = 1cm of E](UE){$U_E$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (A) -- (E);
\draw[->] (UA) -- (A);
\draw[->] (UE) -- (E);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (E) -- (C);
\end{tikzpicture}
```

@@

#

>>>>

# Definition & Assumptions SCM

- Noise terms independent $(N_c,N_e)$
- Mechanisms independent (other variables invariant)(local changes)

Definition 1: A structural causal model M is given by a set of variables $X1, ..., Xd$ and corresponding assignments of the form $X_i := f_i(Pi, Ui), i = 1, ..., d .$
Here, $P_i \subseteq f_{X1}, ..., X_{dg}$ is a subset of the variables that we call the parents of $Xi$. The random variables $U1, ..., Ud$ are called noise variables, which we require to be jointly independent.
The directed graph corresponding to the model has one node for each variable Xi, which has incoming edges from all the parents Pi. 
We will call such a graph the causal graph corresponding to the structural causal model.

@@

\begin{columns}
\begin{column}{0.48\textwidth}

```{tikz,fig.cap="Structural Causal Model", fig.align="center", echo =F}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
  \begin{tikzpicture}[
    sharp corners=2pt,
    inner sep=6pt,
    node distance=2cm,
    >=latex]
\tikzstyle{my node}=[draw, shape = circle, minimum height=1cm,minimum width=1cm]
\tikzstyle{latent}=[draw, shape = rectangle, minimum height=1cm,minimum width=1cm]
\node[my node, fill=gray!30] (A){A};
\node[latent,left of=A](UA){$U_A$};
\node[my node, fill=gray!30] at ($(A)!0.5!(A)-(0pt,1.5cm)$) (E) {E};
\node[my node,right = 1 cm of E, fill=gray!30](C){C};
\node[latent,left = 1cm of E](UE){$U_E$};
\node[latent,right = 1cm of A](UC){$U_C$};
\draw[->] (UA) -- (A);
\draw[->] (UE) -- (E);
\draw[->] (UC) -- (C);
\draw[->] (A) -- (C);
\draw[->] (A) -- (E);
\draw[->] (E) -- (C);
\end{tikzpicture}
```

\end{column}
\begin{column}{0.48\textwidth}
$\{U_A\}$ - Numeric Ability

$\{U_E\}$ - Literary Ability

$\{U_C\}$ - Quality of the Student

$\{A\}$ - Math Test

$\{E\}$ - English-Test

$\{C\}$ - GPA
\end{column}
\end{columns}

@@

# Questions for Discussion

1. How do you think causal model will impact machine learning in the upcoming future? 
2.
